\section{Message from the General Chair}\vspace{2em}
\setheaders%
    {Message from the General Chair}%
    {Message from the General Chair}
\thispagestyle{emptyheader}

\setlength{\parskip}{1ex}

Welcome to Berlin, welcome to ACL-2016!

A special welcome if this is your first ACL -  I hope it fulfills your expectations and remains in your memory as a great start. Its magnitude may be a bit overwhelming - our field is on an expanding trajectory, and even a selection of the best work fills a great number of parallel sessions over a number of days; plus, there are the workshops to quench your topical thirst. Just dive in and drink it up.

If it is your first ACL, then take a good look occasionally at the people in your sessions, listening to your talk or grilling you at your poster; many of these people you will get to know over the years as you grow with them and go to the same meetings in the future. The ACL is not only a breeding ground for project ideas and future collaborations, but is also a hunting ground where job seekers find potential employers. Our new Recruitment Lunch was set up to foster this. Last but not least, ACLs have proven to be the starting moment for lasting friendships. Make the most of it!

To all others, welcome back to another great edition of ACL. The programme chairs and the chair teams for the demos, tutorials, and workshops have put together an exciting programme; the local chairs have fitted this all masterfully in Humboldt University’s buildings at Unter den Linden in the very heart of Berlin. Cast in a quote in the marble entry hall, Karl Marx encourages us to use our insights and technologies to make the world a better place. This tantalizing encouragement is worth pondering over.

On behalf of the entire ACL-2016 organizing team, welcome to Berlin; I wish you a fruitful conference!

\vskip 0.2in
\noindent Antal van den Bosch\\
General Chair

\index{Bosch, Antal}
